---
title: "Clustering"
output: html_document
  latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## Clustering

These examples will include: 

```{r players}
two_players = data.frame(x=c(5, 15),y=c(4, 10))
# Split the players data frame into two observations
player1 <- two_players[1,]
player2 <- two_players[2,]
# Calculate and print their distance using the Euclidean Distance formula
player_distance <- sqrt( (player1$x - player2$x)^2 + (player1$y - player2$y)^2 )
```

## Including Plots

You can also embed plots, for example:

```{r graph, echo=FALSE}
# Plot the positions of the players
ggplot(two_players, aes(x = x, y = y)) + 
  geom_point() +
  # Assuming a 40x60 field
  lims(x = c(-30,30), y = c(-20, 20))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
three_players = data.frame(x = c(5, 15, 0), y = c(4, 10, 20))
#dict() function used in r to calculate euclidean distance
# Calculate the Distance Between two_players
dist_two_players <- dist(two_players)
dist_two_players

# Calculate the Distance Between three_players
dist_three_players <- dist(three_players)
dist_three_players
```
Starndardization I normalize my values and put them in the same scale

Height(scaled) = height - mean(height)
                ----------------------
                      sd(height)
                      
```{r}
three_trees = data.frame(Girth = c(8.3, 8.6, 10.5), Height = c(840, 780, 864))
three_trees

# Calculate distance for three_trees 
dist_trees <- dist(three_trees)

# Scale three trees & calculate the distance  
scaled_three_trees <- scale(three_trees)
dist_scaled_trees <- dist(scaled_three_trees)

# Output the results of both Matrices
print('Without Scaling')
dist_trees

print('With Scaling')
dist_scaled_trees
```

Jaccard index to get measurement for categorical data
J(a,b) = A intersect B/A union B
Distance(a,b) = 1 - J(a,b)

```{r}
job_satisfaction = c("Low", "Low", "Hi", "Low", "Mid")
is_happy = c("No", "No", "Yes", "No", "No")
job_survey = data.frame(job_satisfaction=job_satisfaction,is_happy=is_happy)
job_survey

library(dummies)

# Dummify the Survey Data
dummy_survey <- dummy.data.frame(job_survey)

# Calculate the Distance, method in distribution is binary 
dist_survey <- dist(dummy_survey, method="binary")

# Print the Original Data
job_survey

# Print the Distance Matrix
dist_survey
```
```{r}
players <- data.frame(x=c(-1,-2,8,7,-12,-15),y=c(1,-3,6,-8,8,0))
players
dist_players <- dist(players,method = 'euclidean')
# Perform the hierarchical clustering using the complete linkage
hc_players <- hclust(dist_players, method = 'complete')
# Calculate the assignment vector with a k of 2
cluster_assignments <- cutree(hc_players,k=2)
cluster_assignments
# Create a new dataframe storing these results
players_clustered <- mutate(players, cluster= cluster_assignments)
players_clustered
ggplot(players_clustered,aes(x=x,y=y,color=factor(cluster))) +
  geom_point()
plot(hclust(dist_players, "complete"))
```

```{r}
lineup = structure(list(x = c(-1, -2, 8, 7, -12, -15, -13, 15, 21, 12, 
-25, 26), y = c(1, -3, 6, -8, 8, 0, -10, 16, 2, -15, 1, 0)), .Names = c("x", 
"y"), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA, 
-12L))
# Prepare the Distance Matrix
dist_players <- dist(lineup)

# Generate hclust for complete, single & average linkage methods
hc_complete <- hclust(dist_players, method="complete")
hc_single <- hclust(dist_players, method="single")
hc_average <- hclust(dist_players, method="average")

# Plot & Label the 3 Dendrograms Side-by-Side
par(mfrow = c(1,3))
plot(hc_complete, main = 'Complete Linkage')
plot(hc_single, main = 'Single Linkage')
plot(hc_average, main = 'Average Linkage')
```

```{r}
#install.packages("dendextend")
library(dendextend)
dist_players <- dist(lineup, method = 'euclidean')
hc_players <- hclust(dist_players, method = "complete")

# Create a dendrogram object from the hclust variable
dend_players <- as.dendrogram(hc_players)

# Plot the dendrogram
plot(dend_players)

# Color brances by cluster formed from the cut at a height of 20 & plot
dend_20 <- color_branches(dend_players, h = 20)

# Plot the dendrogram with clusters colored below height 20
plot(dend_20)

# Color brances by cluster formed from the cut at a height of 40 & plot
dend_40 <- color_branches(dend_players, h = 40)

# Plot the dendrogram with clusters colored below height 40
plot(dend_40)

```
```{r}
# Buildin kmeans model
model_km2 <- kmeans(lineup, centers = 2)
# Extract the cluster assignment vector from the kmeans model
clust_km2 <- model_km2$cluster

# Create a new dataframe appending the cluster assignment
lineup_km2 <- mutate(lineup, cluster = clust_km2)

# Plot the positions of the players and color them using their cluster
ggplot(lineup_km2, aes(x = x, y = y, color = factor(cluster))) +
  geom_point()
```

```{r}
#finding k with elbow graph
library(purrr)

# Use map_dbl to run many models with varying value of k (centers)
tot_withinss <- map_dbl(1:10,  function(k){
  model <- kmeans(x = lineup, centers = k)
  model$tot.withinss
})

# Generate a data frame containing both k and tot_withinss
elbow_df <- data.frame(
  k = 1:10 ,
  tot_withinss = tot_withinss
)
# Plot the elbow plot
ggplot(elbow_df, aes(x = k, y = tot_withinss)) +
  geom_line() +
  scale_x_continuous(breaks = 1:10)
```
```{r}
#silhoutte analysis
library(cluster)
#  -1           0           1
# belongs to   undecided    close to
# another                   well match
# cluster
# Generate a k-means model using the pam() function with a k = 2
pam_k2 <- pam(lineup, k = 2)

# Plot the silhouette visual for the pam_k2 model
plot(silhouette(pam_k2))

# Generate a k-means model using the pam() function with a k = 3
pam_k3 <- pam(lineup, k = 3)

# Plot the silhouette visual for the pam_k3 model
plot(silhouette(pam_k3))
```

